Structured Versus Decorated Cospans

### 2021-03-15 cospans_new.tex
started new version "cospans_new.tex"

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/1/22 - first version to Trans. AMS

@@

% MISCELLANEOUS SYMBOLS
\newcommand{\inv}{^{-1}}
\newcommand{\op}{^{\mathit{op}}}
\newcommand{\co}{^{\mathit{co}}}
\newcommand{\coop}{^{\mathit{coop}}}
\newcommand{\id}{\mathrm{id}}
\let\adj\dashv
\newcommand{\pullbackcorner}[1][dr]{\save*!/#1-1.2pc/#1:(-1,1)@^{|-}\restore}
\let\iso\cong
\let\eqv\simeq
\let\cng\equiv
\alwaysmath{ell}
\alwaysmath{infty}
\alwaysmath{odot}
\def\frc#1/#2.{\frac{#1}{#2}}   % \frc x^2+1 / x^2-1 .
\mdef\ten{\mathrel{\otimes}}
\newcommand{\simrightarrow}{\xrightarrow{\raisebox{-3pt}[0pt][0pt]{\ensuremath{\sim}}}}

@@

\newcommand{\double}[1]{\mathbf{\mathbb #1}}
\newcommand{\lCsp}{\double{Csp}}
\newcommand{\lCospan}{\double{Cospan}}
\newcommand{\lOpen}{\double{Open}}
\newcommand{\lA}{\double{A}}
\newcommand{\lB}{\double{B}}
\newcommand{\lC}{\double{C}}
\newcommand{\lD}{\double{D}}
\newcommand{\lE}{\double{E}}
\newcommand{\lF}{\double{F}}
\newcommand{\lG}{\double{G}}
\newcommand{\lH}{\double{H}}
\newcommand{\lR}{\double{R}}
\newcommand{\lX}{\double{X}}
\newcommand{\lY}{\double{Y}}

@@

Here we present a solution to this problem.  Instead of basing the theory of decorated cospans on a symmetric lax monoidal functor $F \maps (\A, +) \to (\Set, \times)$, we use a symmetric lax monoidal pseudofunctor $F \maps (\A, +) \to (\Cat, \times)$.  In \cref{thm:decorated_cospans,DC}, we use this
data to construct a symmetric monoidal double category $F\lCsp$ in which:
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
\node (D) at (3,0) {$x \in F(m)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a \define{map of decorated cospans}: that is, a commutative
diagram
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a morphism $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}

@@

We thank Daniel Cicala, Brendan Fong, and Joe Moeller for helpful conversations. The third author would  like  to  thank  the  General  Secretariat  for  Research  and Technology (GSRT) and the Hellenic Foundation for Research and Innovation (HFRI).

@@

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction. The definition of a lax monoidal pseudofunctor is recalled in \cref{subsec:bicats}, and the definition of a double category is recalled in \cref{defn:double_category}.

@@

\begin{thm}\label{thm:decorated_cospans}
Let $\A$ be a category with finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\begin{equation}\label{eq:mapofdec}
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{decoration morphism} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

\begin{proof}
Take $\A$ to have finite colimits and let $(F,\phi,\phi_0) \maps (A,+) \to (\Cat,\times)$ be a lax monoidal pseudofunctor.  We shall define the category of objects $F\lCsp_0$, the category of arrows $F\lCsp_1$, as well as the four structure functors which tie them together: the source and target functors $S,T \maps F\lCsp_1 \to F\lCsp_0$, the unit functor $U \maps F\lCsp_0 \to F\lCsp_1$, and the composition functor $\odot \maps F\lCsp_1 \times_{F\lCsp_0} F\lCsp_1 \to F\lCsp_1$.

The category of objects $F\lCsp_0$ is simply $\A$, whereas $F$-decorated cospans and maps between them form the category of arrows $F\lCsp_1$. The (vertical) composition of two maps of $F$-decorated cospan is 
\begin{equation}\label{eq:verticalcompo}
 \begin{tikzcd}
a\ar[r,"i"]\ar[d,"f"'] & m\ar[d,"h"] & b\ar[l,"o"']\ar[d,"g"]  & x\in F(m) \\
a'\ar[r,"{i'}"]\ar[d,"{f'}"'] & m'\ar[d,"{h'}"] & b'\ar[l,"{o'}"']\ar[d,"{g'}"] & x'\in F(m') \\
a''\ar[r,"{i''}"] & m'' & b''\ar[l,"{o''}"'] & x''\in F(m'')
 \end{tikzcd}\;=\;
 \begin{tikzcd}[column sep=.3in]
a\ar[r,"i"]\ar[dd,"{f'f}"'] & m\ar[dd,"{h'h}"] & b\ar[l,"o"']\ar[dd,"{g'g}"]  & x\in F(m) \\
\hole \\
a''\ar[r,"{i''}"] & m'' & b''\ar[l,"{o''}"'] & x''\in F(m'')  
 \end{tikzcd}
\end{equation}
where if $\tau \maps F(h)(x) \to x'$ and $\tau' \maps F(h')(x') \to x''$ are the decoration morphisms, 
\begin{equation}\label{eq:tauofverticalcomposite}
F(h'h)(x) \simrightarrow F(h')F(h)(x)\xrightarrow{F(h')(\tau)}F(h')(x')\xrightarrow{\tau'}x''
\end{equation}
is the decoration of the composite, where the first isomorphism comes from pseudofunctoriality of $F$.

The source and target functors $S$ and $T$ map an $F$-decorated cospan to the source and target of the cospan and a map of decorated cospans \cref{eq:mapofdec} to $f$ and $g$ respectively. The unit functor $U$ maps an object $a$ to 
\begin{equation}\label{eq:UFCsp}
\begin{tikzcd}
a\ar[r,"1_a"] & a & a,\ar[l,"1_a"'] & \bot_a:=\one\xrightarrow{\phi_0}F(0)\xrightarrow{F(!_a)}F(a)
\end{tikzcd}
\end{equation}
where $\bot_a \in F(a)$ is called the \define{trivial decoration}, and a vertical 1-morphism $f\maps a\to a'$ to
\node (B) at (1,0.5) {$a$};
\node (C) at (2,0.5) {$a$};
\node (C') at (2,-0.5) {$a'$};
\node (D) at (1,-0.5) {$a'$};
\node (E) at (3,0.5) {$\bot_a \in F(a)$};
\node (F) at (3,-0.5) {$\bot_{a'} \in F(a')$};
(A) edge node[above]{$1_a$} (B)
(C) edge node[above]{$1_a$} (B)
(C) edge node[right]{$f$} (C')
(A') edge node[above] {$1_{a'}$} (D)
(C') edge node[above] {$1_{a'}$} (D)
(B) edge node [left] {$f$} (D);
together with the decoration morphism $F(f)F(!_a)\phi_0(*)\cong F(!_{a'})\phi_0(*)$.
Finally, the (horizontal) composition functor $\odot$ maps two composable $F$-decorated cospans 
\[  M=\left(a\to m\leftarrow b,x\in F(m)\right), \qquad N=\left(b\to n\leftarrow c,y\in F(n)\right)\] 
to $N\odot M$, which as a cospan is their pushout in $\A$ over their shared foot, shown in dashed arrows here:
\begin{equation}\label{eq:dcospanscomposition}
\begin{tikzpicture}[scale=1.2,baseline=(current bounding box.center)]
\node (B) at (1,1) {$m$};
\node (D) at (3,1) {$n$};
\node (E) at (4,0) {$c$};
\node (F) at (2,2) {$m+n$};
\node (G) at (2,3) {$m+_{b} n$};
(C) edge node [above] {$i'$} (D)
(E) edge node [above] {$o'$} (D)
(B) edge node [below] {$\iota_m$} (F)
(D) edge node [below] {$\iota_{n}$} (F)
(B) edge[bend left=10] node [above] {$v_m\;$} (G)
(D) edge[bend right=10] node [above] {$\;v_n$} (G)
(F) edge node [fill=white] {$\psi$} (G)
(A) edge[dashed,bend left] node [left] {} (G)
(E) edge[dashed,bend right] node [right] {} (G);
\end{equation}
equipped with the decoration
\[  y\odot x:= \one\cong\one \times \one \xrightarrow{x \times y} F(m) \times F(n) \xrightarrow{\phi_{m,n}} F(m+n) \xrightarrow{F(\psi)} F(m+_{b}n) \]
where $\psi \maps m + n \to m+_{b} n$ is the natural map from the coproduct to the pushout. Moreover, given two horizontally composable maps of $F$-decorated cospans $\alpha$ and $\beta$ 
\node (D) at (1,-0.5) {$m'$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\node (G) at (4,0.5) {$b$};
\node (H) at (5,0.5) {$n$};
\node (I) at (6,0.5) {$c$};
\node (G') at (4,-0.5) {$b'$};
\node (H') at (5,-0.5) {$n'$};
\node (I') at (6,-0.5) {$c'$};
\node (J) at (7,0.5) {$y \in F(n)$};
\node (K) at (7,-0.5) {$y' \in F(n')$};
\node (L) at (1,-1) {$\tau_\alpha \maps F(h_1)(x) \to x'$};
\node (M) at (5,-1) {$\tau_\beta \maps F(h_2)(y) \to y'$};
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A') edge node[above] {$i_1'$} (D)
(C') edge node[above] {$o_1'$} (D)
(B) edge node [left] {$h_1$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$g$} (G')
(H) edge node [left] {$h_2$} (H')
(G') edge node [above] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$k$} (I')
(I') edge node [above] {$o_2'$} (H');
their composite $\beta\odot \alpha$ is the horizontal composite of the two cospans in $\A$
\begin{equation}\label{eq:horizontalcompo}
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (B) at (1.5,0.5) {$m+_{b} n$};
\node (C) at (3,0.5) {$c$};
\node (C') at (3,-0.5) {$c'$};
\node (D) at (1.5,-0.5) {$m'+_{b'}n'$};
\node (E) at (4.5,0.5) {$y \odot x \in F(m+_b n)$};
\node (F) at (4.5,-0.5) {$y' \odot x' \in F(m' +_{b'} n')$};
(A) edge node[above]{} (B)
(C) edge node[above]{} (B)
(C) edge node[right]{$k$} (C')
(A') edge node [above]{} (D)
(C') edge node [above]{} (D)
(B) edge node [left] {$h_1 +_g h_2$} (D);
\end{equation}
together with the decoration morphism $\tau_{\beta \odot \alpha} \maps F(h_1 +_g h_2)(y \odot x) \to (y' \odot x')$ given by the diagram:
\begin{equation}\label{eq:decohorizontalcompo}
\begin{tikzpicture}[scale=1.7,baseline=(current bounding box.center)]
\node (A) at (4.75,0) {$\scriptstyle\Downarrow \tau_\alpha \times \tau_\beta$};
\node (D) at (3,0) {$\one \cong \one \times \one$};
\node (E) at (5.5,0.5) {$F(m) \times F(n)$};
\node () at (6.5,0) {$\stackrel{\phi_{h_1,h_2}}{\cong}$};
\node () at (8.5,0) {$\cong$};
\node (E') at (5.5,-0.5) {$F(m') \times F(n')$};
\node (B) at (7.5,0.5) {$F(m+n)$};
\node (B') at (7.5,-0.5) {$F(m' + n')$};
\node (C) at (9.5,0.5) {$F(m+_{b} n)$};
\node (C') at (9.5,-0.5) {$F(m' +_{b'} n')$};
(E) edge node [above] {$\phi_{m,n}$} (B)
(E') edge node [above] {$\phi_{m',n'}$} (B')
(B) edge node [above] {$F(\psi)$} (C)
(B') edge node [above] {$F(\psi)$} (C')
(C) edge node  [fill=white] {$F(h_1 +_g h_2)$} (C')
(B) edge node [fill=white] {$F(h_1 + h_2)$} (B')
(D) edge node [above] {$x \times y$} (E)
(D) edge node [below] {$x' \times y'$} (E')
(E) edge node [fill=white] {$F(h_1) \times F(h_2)$} (E');
\end{equation}
where the middle isomorphism is \cref{eq:pseudonaturality} from pseudonaturality of $\phi$ and the right-hand side isomorphism is due to pseudofunctoriality of $F$.

>>>>

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/3/15 - attempt at new version

@@

% MISCELLANEOUS SYMBOLS
\newcommand{\inv}{^{-1}}
\newcommand{\op}{^{\mathrm{op}}}
\newcommand{\co}{^{\mathrm{co}}}
\newcommand{\coop}{^{\mathit{coop}}}
\newcommand{\id}{\mathrm{id}}
\let\adj\dashv
\newcommand{\pullbackcorner}[1][dr]{\save*!/#1-1.2pc/#1:(-1,1)@^{|-}\restore}
\let\iso\cong
\let\eqv\simeq
\let\cng\equiv
\alwaysmath{ell}
\alwaysmath{infty}
\alwaysmath{odot}
\def\frc#1/#2.{\frac{#1}{#2}}   % \frc x^2+1 / x^2-1 .
\mdef\ten{\mathrel{\otimes}}
\newcommand{\simrightarrow}{\xrightarrow{\raisebox{-3pt}[0pt][0pt]{\ensuremath{\sim}}}}

@@

\newcommand{\double}[1]{\mathbf{\mathbb #1}}
\newcommand{\lCsp}{\double{Csp}}
\newcommand{\lCospan}{\double{Cospan}}
\newcommand{\lOpen}{\double{Open}}
\newcommand{\Fr}{\double{Fr}}
\newcommand{\lA}{\double{A}}
\newcommand{\lB}{\double{B}}
\newcommand{\lC}{\double{C}}
\newcommand{\lD}{\double{D}}
\newcommand{\lE}{\double{E}}
\newcommand{\lF}{\double{F}}
\newcommand{\lG}{\double{G}}
\newcommand{\lH}{\double{H}}
\newcommand{\lR}{\double{R}}
\newcommand{\lX}{\double{X}}
\newcommand{\lY}{\double{Y}}

@@

Here we present a solution to this problem.  Instead of basing the theory of decorated cospans on a symmetric lax monoidal functor $F \maps (\A, +) \to (\Set, \times)$, we use a symmetric lax monoidal pseudofunctor $F \maps (\A, +) \to (\Cat, \times)$.  In \cref{thm:decorated_cospans,thm:decorated_cospans_2}, we use this
data to construct a symmetric monoidal double category $F\lCsp$ in which:
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
\node (D) at (3,0) {$x \in F(m)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a \define{map of decorated cospans}: that is, a commutative
diagram
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a morphism $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}

@@

We thank Daniel Cicala, Brendan Fong, Joe Moeller and especially Michael Shulman for helpful conversations. The third author would like to thank the General Secretariat for Research  and Technology (GSRT) and the Hellenic Foundation for Research and Innovation (HFRI).

@@

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction.   \cref{thm:decorated_cospans} explains how to construct a double category of decorated cospans from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$  whenever $A$ has finite colimits ({\bf or is just cocartesian?}).  \cref{thm:decorated_cospans_2} gives conditions under which this double category is symmetric monoidal.   The definition of lax monoidal pseudofunctor is recalled in \cref{subsec:bicats}, and the definition of symmetric monoidal double category is recalled in \cref{defn:double_category}.

@@

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\begin{equation}\label{eq:mapofdec}
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{decoration morphism} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

\begin{thm}\label{thm:decorated_cospans_2}
Let $\A$ be a category with finite colimits and let $(F,\phi,\phi_0)\maps (\A,+) \to (\Cat,\times)$ be a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ is symmetric monoidal, where the tensor product
\begin{itemize}
\item of two objects $a$ and $b$ is their coproduct $a+b$ in $\A$,
\item of two vertical 1-morphisms $f \maps a \to b$ and $f' \maps a' \to b'$ is $f+f' \maps a+a' \to b+b'$ in $\A$,
\item of two horizontal 1-cells $(a \to m \leftarrow b ,x \in F(m))$ and $(c \to n \leftarrow d, y\in F(n))$ is
\begin{equation}\label{eq:tensordecoration}
\begin{tikzcd}
& m+n & \\
a+c\ar[ur,"i_1+i_2"] && b+d,\ar[ul,"o_1+o_2"']
\end{tikzcd}\qquad x \otimes y:=  \phi_{m,n}(x,y) \in F(m + n) 
\end{equation}
\item of two 2-morphisms $\alpha$ and $\beta$ is:
\[
\begin{tikzpicture}[scale=1.2]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
%\node (E) at (2.7,0.5) {$\scriptstyle x \in F(m)$};
\node () at (2.7,0) {$\otimes$};
\node () at (6.2,0) {$=$};
%\node (F) at (2.7,-0.5) {$\scriptstyle x' \in F(m')$};
\node (G) at (3.5,0.5) {$c$};
\node (H) at (4.5,0.5) {$n$};
\node (I) at (5.5,0.5) {$d$};
\node (G') at (3.5,-0.5) {$c'$};
\node (H') at (4.5,-0.5) {$n'$};
\node (I') at (5.5,-0.5) {$d'$};
%\node (J) at (6.7,0.5) {$\scriptstyle y \in F(n)$};
%\node (K) at (6.7,-0.5) {$\scriptstyle y' \in F(n')$};
\node (L) at (1,-1.2) {$\scriptstyle\tau_\alpha \maps F(h)(x) \to x'\textrm{ in } F(m')$};
\node (M) at (4.5,-1.2) {$\scriptstyle\tau_\beta \maps F(h')(y) \to y'\textrm{ in } F(n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i_1'$} (D)
(C') edge node[below] {$o_1'$} (D)
(B) edge node [left] {$h$} (D)
(G) edge node [above] {$i_2$} (H)
(G) edge node [left] {$f'$} (G')
(H) edge node [left] {$h'$} (H')
(G') edge node [below] {$i_2'$} (H')
(I) edge node [above] {$o_2$} (H)
(I) edge node [right] {$g'$} (I')
(I') edge node [below] {$o_2'$} (H');
\end{tikzpicture}
\begin{tikzpicture}[scale=1.1]
\node (A) at (0,0.5) {$a+c$};
\node (A') at (0,-0.5) {$a'+c'$};
\node (B) at (2,0.5) {$m+n$};
\node (C) at (4,0.5) {$b+d$};
\node (C') at (4,-0.5) {$b'+d'$};
\node (D) at (2,-0.5) {$m'+n'$};
%\node (E) at (5.2,0.5) {$\scriptstyle \phi_{m,n}(x,y)$};
%\node (F) at (5.2,-0.5) {$\scriptstyle \phi_{m',n'}(x',y')$};
\node () at (2,-1.3) {$\scriptstyle\tau_{\alpha\ot\beta}\maps F(h+h')(\phi_{m,n}(x,y))\to \phi_{m',n'}(x',y')\textrm{ in }F(m'+n')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B)
(A) edge node[left]{$f+f'$} (A')
(C) edge node[right]{$g+g'$} (C')
(A') edge node [below]{$i_1'+i_2'$} (D)
(C') edge node [below]{$o_1'+o_2'$} (D)
(B) edge node [left] {$h+h'$} (D);
\end{tikzpicture}
\]
with decoration morphism $\tau_{\alpha\otimes\beta}$ given by the following diagram:
\begin{displaymath}
 \begin{tikzcd}[column sep=.5in, row sep=.1in]
& F(m)\times F(n)\ar[r,"\phi_{m,n}"]\ar[dd,"F(h)\times F(h')"description]\ar[ddr,phantom,"{\stackrel[\cref{eq:pseudonaturality}]{\phi_{h,h'}}{\cong}}"] & F(m+n)\ar[dd,"F(h+h')"] \\
\one \ar[ur,"x\times y"]\ar[dr,"x'\times y'"']\ar[r,phantom,"\scriptstyle\Downarrow\tau_\alpha\times\tau_\beta"] & \phantom{A} && \\
& F(m')\times F(n')\ar[r,"\phi_{m',n'}"'] & F(m'+n')
 \end{tikzcd}
\end{displaymath}
\end{itemize}
\end{thm}

We prove both these theorems using the work of Shulman \cite{Shulman2008}, which we recall here.   First we need a few concepts.   Opfibrations are recalled in \cref{subsec:fibrations}, 
and fibrations are defined dually, with a functor $\Phi \maps \C \to \A$ being a fibration if and only
if $\Phi\op \maps \C\op \to \A\op$ is an opfibration.   A functor $\Phi \maps \C \to \A$ is a \define{bifibration} if it is both a fibration and an opfibration.  If $\Phi \maps \C \to \A$ an opfibration we have for any morphism $f \maps a \to b$ in $\A$ a covariant reindexing functor
\[    f_! \maps \C_a \to \C_b \]
between the fibers of $\Phi$, as recalled in \cref{subsec:fibrations}.  Similarly, if $\Phi$ is a fibration we have a contravariant reindexing functor
\[    f^\ast \maps \C_b \to \C_a .\]
Thus, in a bifibration we have both covariant and contravariant reindexing functors, and in 
fact $f^\ast$ is left adjoint to $f_\ast$ \cite[Proposition 3.9]{Shulman2008}.   Using this one can easily show that for any commutative square in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1,-1) {$d$};
%\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
the following square commutes up to a natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\C_a$};
\node (B) at (1,0) {$\C_b$};
\node (C) at (0,-1) {$\C_c$};
\node (D) at (1,-1) {$\C_d$};
\node (E) at (0.5,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^\ast$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^\ast$} (C);
\end{tikzpicture}
\]
If $\theta$ is a natural isomorphism whenever the original square in $\A$ was
a pushout, we say $\Phi$ is \define{Beck--Chevalley}.

\begin{lem}[\textbf{Shulman}] \label{lem:shulman}
Suppose $\A$ has finite colimits and $\Phi \maps \C \to \A$ is a Beck--Chevalley bifibration.  Then there is a double category $\Fr(\Phi)$ in which

\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is a pair of objects $a,b \in \A$ together with an
object $c \in C$ with $\Phi(c) = a + b$.
\item a 2-morphism is a pair of morphisms $f \maps a \to a'$, $g \maps b \to b'$ 
in $\A$ together with a morphism $h \maps c \to c'$ with $\Phi(h) = f + g$.
\end{itemize}
If $\Phi$ is a symmetric monoidal bifibration then $\Fr(\Phi)$ becomes a symmetric
monoidal double category.
\end{lem}

\begin{proof} This is \cite[Theorem 14.9]{Shulman2008}; Shulman proves the dual version in more detail in his Theorem 14.4.  
\end{proof}

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] We apply \cref{lem:shulman} to ....
\end{proof}

\begin{proof}[Proof of \cref{thm:decorated_cospans_2}] We apply \cref{lem:shulman} to ....
\end{proof}

\begin{comment}
We can draw the 2-morphism as follows, {\bf but with ticks on the horizontal arrows!}:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1,-1) {$b'$};
\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$c$} (B)
(A) edge node[left]{$f$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[above]{$c'$} (D);
\end{tikzpicture}
\]
\end{comment}
